%!TEX encoding = UTF-8 Unicode

%% Thesis cover pages
%%
%% Based on the example.tex file provided by Tomás Oliveira e Silva
%%
%% Modified by João Paulo Barraca <jpbarraca@ua.pt>
%% Modified by Filipe Manco <filipe.manco@ua.pt>

% Remove 'coverdepname' parameter to remove department name from cover pages
% Add 'final' parameter to remove "DOCUMENTO PROVISÓRIO" from cover page
\documentclass[11pt,a4paper]{report}

%%% This file is a template for the cover of your
%%%  thesis.
%%% Lines like this one, starting with %%% , are
%%%  instructions. Search them to know where and
%%%  how to edit this file. After editing the
%%%  document you can delete them.
%%% Items marked with OPTIONAL can be removed.
%%% Most of the lines you will have to change
%%%  already have examples to help you understand.
%%% The complete list of options and commands can
%%%  be found in uaThesisTemplate.sty

%%% Select your area of study and department from
%%%  the following lists and replace them in the
%%%  \usepackage declaration ([area,department]).
%%%
%%% | area        |   | department |
%%% ---------------   --------------
%%% | arts        |   | dao        |
%%% | sciences    |   | dbio       |
%%% | education   |   | dcspt      |
%%% | economy     |   | deca       |
%%% | engineering |   | degeit     |
%%% | humanities  |   | de         |
%%% | health      |   | deti       |
%%% |             |   | demac      |
%%% |             |   | decivil    |
%%% |             |   | dem        |
%%% |             |   | dfis        |
%%% |             |   | dlc        |
%%% |             |   | dmat       |
%%% |             |   | dq         |
%%% |             |   | dcm        |
%%% |             |   | dgeo        |
%%% ---------------   --------------
%% Import the UA thesis template
\usepackage[engineering,deti]{uaThesisTemplate}

%% Optional packages for utf8 encoding and T1 fonts
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\renewcommand\familydefault{\sfdefault}

%% We will use english and portuguese
\usepackage[english,portuguese]{babel}

\usepackage{xspace}% used by \sigla

%%% This package is only used to create a dummy
%%%  text for the abstract. You can remove the
%%%  following line in the final document.
%\usepackage{blindtext}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%       DEFINE THESIS COVER FIELDS
%%                  START
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% Edit thesis year
%% Thesis Year
\ThesisYear{\the\year}


%%% Uncomment one of the following options:
%\DocumentTypeThesis
\DocumentTypeThesisProposal
%\DocumentTypeDissertation

%%% Uncomment if you are a student of a MAP Doctoral program 
%%% http://www.map.edu.pt/
%%% | area                | identifier |
%%% | Biology of Plants   | bioplant   |
%%% | Computer Science    | i          |
%%% | Physics             | fis        |
%%% | Telecommunications  | tele       |

%\MAPProgram{i}

%%% Edit author's name
%% Author's full name
\Author{Pedro\\Azevedo}

%%% Edit thesis title in Portuguese
%% Thesis Mandatory Title in Portuguese
\TitlePT{The Playground: Plataforma móvel para conectar jogadores casuais de basquetebol}

%%% OPTIONAL: you can write your thesis title in
%%%  other languages. You can repeat the following
%%%  line as many times as you want, or remove it
%%%  if you only want the portuguese title.
%% Optional Titles in additional languages.
\TitleEN{The Playground: mobile platform to connect amateur basketball players}

%%% Edit your supervisor's information.
%%%
%%% Use this command for male supervisor
%%% \Supervisor{name}{category}
%%%
%%% Or this command for female supervisor
%%% \Supervisora{name}{category}
%%%
%%% name   : Supervisor's name
%%% categoy: One of the following list
%%%
%%% | category                  |
%%% -----------------------------
%%% | catedrático/a             |
%%% | catedrático/a convidado/a |
%%% | associado/a c/ agregação  |
%%% | associado/a               |
%%% | auxiliar                  |
%%% | auxiliar convidado        |
%%% -----------------------------
%% Supervisor information
\Supervisor{Ilídio Oliveira}{associado}

%%% OPTIONAL: you can add a second supervisor.
%%%
%%% Use this command for male supervisor
%%% \SupervisorMember{name}{category}
%%% 
%%% Or this command for female supervisor
%%% \SupervisoraMember{name}{category}
%%%  {[preposition: (at, in, no, da)] department}
%%%  {[preposition: (at, in, do, na)] school}
%%%  {[preposition: (and, e)] }
%%%
%%% name      : See above.
%%% category  : See above.
%%% department: Can be a department, faculty or
%%%             other; Choose the preposition
%%%             accordingly.
%%% school    : University/School name; Choose the
%%%             preposition accordingly.
%%% last      : Write {e } (Portuguese) or {and } (English)
%%%             for the last enumerated Supervisor; Or
%%%             nothing {} for all other supervisors 
%% Optional SecondSupervisor information
%%\SupervisorMember{(co-orientador)}{auxiliar convidado}{do Departamento de Matemática}{da Universidade de Aveiro}{}

%%% Repeat the above declaration for additional
%%%  Supervisor members as many times as necessary.
%% Additional Supervisor members
%%\SupervisoraMember{(co-orientadora)}{associada c/ agregação}{do Departamento de Biologia}{da Universidade de Aveiro}{}
%%\SupervisorMember{(co-orientador)}{auxiliar convidado}{do Departamento de Física}{da Universidade de Aveiro}{e }

%%% Edit your degree information.
%%%
%%% \Degree{title}{name}
%%%
%%% title: Degree title (Mestre or Doutor)
%%% name : Degree name
%% Degree title and name
\Degree{Mestre}{Engenharia Informática}

%%% OPTIONAL: edit grant support texts where
%%%  appropriate. Maximum two.
%%%
%%% \GrantText{grant1}{grant2}
%% Optional Grant support texts
%\GrantText{Texto Apoio financeiro do POCTI no âmbito do III Quadro Comunitário de Apoio.\\
%          \newline
%          (if applicable- Portuguese mandatory)}
%          {Texto Apoio financeiro da FCT e do FSE no âmbito do III Quadro Comunitário de Apoio.\\
%          \newline
%          (if applicable- Portuguese mandatory)}

%%% Edit the Jury President information.
%% First Member will be the President of the Jury
%\JuriMember{Prof. Doutor João Antunes da Silva}{professor associado da Universidade de Aveiro}

%%% Repeat the above declaration for additional
%%%  members of the jury as many times as necessary.
%% Additional members of the Jury
%\JuriMember{Prof. Doutor João Antunes da Silva}{professor associado da Universidade de Aveiro}
%\JuriMember{Prof. Doutor João Antunes da Silva}{professor associado da Universidade de Aveiro}
%\JuriMember{Prof. Doutor João Antunes da Silva}{professor associado da Universidade de Aveiro}
%\JuriMember{Prof. Doutor João Antunes da Silva}{professor associado da Universidade de Aveiro}
%\JuriMember{Prof. Doutor João Antunes da Silva}{professor associado da Universidade de Aveiro}



%%% OPTIONAL: Add an acknowledgement text
%% Optional Acknowledgement text
%\Acknowledgements{Agradeço toda a ajuda a todos os meus colegas e companheiros.}

%%% Edit the keywords and abstract in Portuguese
%%% Don't forget to remove the \blindtext[3]
%%%  declaration.
%%% Don't remove the \selectlanguage statement or you will end with incorrect hyphenation
%% Mandatory Keywords and Abstract in Portuguese
\Abstract{palavras-chave}{Basquetebol, Desporto Casual, Desporto Amador, Kotlin Multiplatform, Desenvolvimento de Aplicações Móveis, Android, iOS, Organização de Jogos, Disponibilidade de Campos}
         {resumo}{\selectlanguage{portuguese} Este projeto de dissertação de mestrado procura produzir uma solução para um problema encontrado em jogadores casuais e amadores de basquetebol que jogam em campos públicos. Vários destes jogadores encontram dificuldades em formar equipas equilibradas: quando vão a um campo público, ou não existem jogadores suficientes, ou quando existem, não estão no mesmo nível de competitividade e experiência. O contrário também acontece: o campo pode estar cheio encontrando dificuldades em partilhar espaço com jogadores que apenas querem lançar bolas. Tomando a perspetiva de um jogador à procura de campos públicos e jogos para participar, foi feita uma pesquisa na internet e em modelos generativos de inteligência artificial. Várias plataformas foram encontradas, mas parecem não ser utilizadas em Portugal. Desta forma, abre-se uma oportunidade de desenvolvimento de uma aplicação para diferentes plataformas móveis focada na comunidade de basquetebol portuguesa. O projeto segue metodologia User-Centred Design nas três fases e Agile Scrum nas fases desenvolvimento com Sprints. Na primeira fase foi desenvolvido um questionário com 50 respostas que confirmou os problemas identificados: 22\% dos jogadores não tem com quem ir jogar, e 10\% menciona falta de jogadores em campo. Quanto aos objetivos, 29\% vai jogar com amigos contra outros jogadores, e 17\% vai apenas lançar bolas. Foram desenvolvidas duas Personas: Miguel (jogador competitivo de 25 anos procurando competição regular) e Carlos (lançador flexível de 33 anos com dificuldade em encontrar espaço para treinar o lançamento). Baseado nas personas, foram recolhidos requisitos em formato de User Stories e divididos pelos Sprints.Optou-se por Kotlin Multiplatform para o desenvolvimento da aplicação e Compose Multiplatform para a interface, funcionando em Android e iOS. Esta proposta estabelece as bases para uma aplicação multiplataforma que responde a uma necessidade genuína na comunidade de basquetebol portuguesa.  Esta proposta estabelece as bases para uma aplicação multiplataforma que responde a uma necessidade genuína na comunidade de basquetebol portuguesa. Jogadores casuais enfrentam barreiras significativas na coordenação de jogos e avaliação da dinâmica dos campos públicos. A solução proposta, que integra localização e ocupação dos campos em tempo real, organização de jogos e sistemas de competição, oferece uma abordagem abrangente a este problema. Embora o sucesso da plataforma dependa criticamente da adoção inicial por parte da comunidade, o processo iterativo adotado através de sprints permite validação contínua com utilizadores reais e ajuste rápido de funcionalidades. Desta forma, a presente dissertação estabelecerá não apenas um produto funcional, mas também uma framework replicável para plataformas comunitárias similares em contextos portugueses.}

\Abstract{keywords}{Basketball, Casual Sports, Amateur Sports, Kotlin Multiplatform, Mobile Application Development, Android, iOS, Game Organization, Court Availability}
         {abstract}{\selectlanguage{english}This master's dissertation project seeks to produce a solution to a problem encountered by casual and amateur basketball players who play on public courts. Several of these players face difficulties in forming balanced teams: when they go to a public court, either there are not enough players, or when there are, they are not at the same level of competitiveness and experience. The opposite also happens: the court can be full, finding difficulties in sharing space with players who just want to shoot balls. Taking the perspective of a player looking for public courts and games to participate in, a search was conducted on the internet and with generative artificial intelligence models. Several platforms were found, but they do not seem to be used in Portugal. Thus, an opportunity opens up for developing an application for different mobile platforms focused on the Portuguese basketball community. The project follows User-Centred Design methodology in all three phases and Agile Scrum in the development phases with Sprints. In the first phase, a questionnaire with 50 responses was developed that confirmed the identified problems: 22\% of players have no one to go play with, and 10\% mention lack of players on the court. Regarding objectives, 29\% go to play with friends against other players, and 17\% go just to shoot balls. Two Personas were developed: Miguel (competitive player, 25 years old seeking regular competition) and Carlos (flexible shooter, 33 years old with difficulty finding space to practice shooting). Based on the personas, requirements were gathered in the format of User Stories and divided by Sprints. Kotlin Multiplatform was chosen for application development and Compose Multiplatform for the interface, working on Android and iOS. This proposal establishes the foundation for a cross-platform application that addresses a genuine need in the Portuguese basketball community. Casual players face significant barriers in coordinating games and evaluating the dynamics of public courts. The proposed solution, which integrates real-time location and occupancy of courts, game organization, and competition systems, offers a comprehensive approach to this problem. Although the success of the platform depends critically on initial adoption by the community, the iterative process adopted through sprints allows continuous validation with real users and rapid adjustment of features. Thus, this dissertation will establish not only a functional product, but also a replicable framework for similar community platforms in Portuguese contexts.}


%%% OPTIONAL: Acknowledge the use of AI Tools
\AIUse{I acknowledge the use of Perplexity AI pro (https://www.perplexity.ai/) to search for documents, rephrasing and enrich text, spot spelling and grammar mistakes and structure tables with the content provided. Also the user of TexStudio (https://www.texstudio.org/) to write the document.}
%I acknowledge the use of [software, codes or platforms] to [specific use software, codes or platforms or to other tasks].

%{Example 1: I acknowledge the use of ChatGPT 3.5 (Open AI, https://chat.openai.com) to summarise the initial notes and to proofread the final draft and the use of
%Office365 (Microsoft, https://www.office.com) for text writing and productivity.}
%\newline\newline
%{Example 2: No content generated by AI technologies has been used in this Thesis.}
%}
%Reconheço o uso de [inserir sistema(s) de IA e respetiva hiperligação] para [indicar utilização específica de IA ou outras tarefas]. Reconheço a utilização de [indicar software, código ou plataforma] para [indicar utilização específica de software, código ou platafoma ou para outras tarefas].
%Exemplo 1: Reconheço a utilização do ChatGPT 3.5 (Open AI, https://chat.openai.com) para resumir as notas iniciais e para rever o rascunho final, a utilização do Office365 (Microsoft, https://www.office.com) para a escrita do texto.
%Exemplo 2: Não foram utilizados no presente trabalho quaisquer conteúdos gerados por tecnologias de IA.



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%       DEFINE THESIS COVER FIELDS
%%                  END
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\begin{document}

%%% By default we consider portuguese to be the
%%%  main language (the last one in the usepackage
%%%  declaration). If you're writing your thesis
%%%  in english uncomment the following line.
%%% Set the proper language or you will end with incorrect hyphenation!
\selectlanguage{english}


\MakeCover
\end{document}
